---
title: "Natural Language Processing"
author: "Puja Roy"
date: "3/31/24"
output: openintro::lab_report
---

### Objective

In Text Mining with R, Chapter 2 looks at Sentiment Analysis.  In this assignment, you should start by getting the primary example code from chapter 2 working in an R Markdown document.  You should provide a citation to this base code.  

You’re then asked to extend the code in two ways:
Work with a different corpus of your choosing, and
Incorporate at least one additional sentiment lexicon (possibly from another R package that you’ve found through research).

Citation: Text Mining with R, Chapter 2: https://www.tidytextmining.com/sentiment.html


### Load Packages
```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

### 2.1 The sentiments datasets
```{r}
#install.packages('tidytext')
#install.packages('textdata')
library(tidytext)

get_sentiments("afinn")
```
```{r}
get_sentiments("bing")
```
```{r}
get_sentiments("nrc")
```

### 2.2 Sentiment analysis with inner join
```{r}
library(janeaustenr)
library(dplyr)
library(stringr)

tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```

```{r}
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```
```{r}
library(tidyr)

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```
```{r}
library(ggplot2)

ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```
### 2.3 Comparing the three sentiment dictionaries
```{r}
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice
```
```{r}
afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```
```{r}
bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```

```{r}
get_sentiments("nrc") %>% 
  filter(sentiment %in% c("positive", "negative")) %>% 
  count(sentiment)
```
```{r}
get_sentiments("bing") %>% 
  count(sentiment)
```
### 2.4 Most common positive and negative words
```{r}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts
```
```{r}
bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```
```{r}
custom_stop_words <- bind_rows(tibble(word = c("miss"),  
                                      lexicon = c("custom")), 
                               stop_words)

custom_stop_words
```

### 2.5 Wordclouds
```{r}
library(wordcloud)

tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```
```{r}
library(reshape2)

tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)
```
### 2.6 Looking at units beyond just words
```{r}
p_and_p_sentences <- tibble(text = prideprejudice) %>% 
  unnest_tokens(sentence, text, token = "sentences")
```

```{r}
p_and_p_sentences$sentence[2]
```
```{r}
austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()

austen_chapters %>% 
  group_by(book) %>% 
  summarise(chapters = n())
```
```{r}
bingnegative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())

tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  slice_max(ratio, n = 1) %>% 
  ungroup()
```

### Selecting a Different Corpus

I selected the corpus: The Beautiful and Damned from the gutenbergr library.
```{r}
library(gutenbergr)
```

```{r}
# Get metadata for all available works in the Gutenberg corpus
gutenberg_metadata <- gutenberg_works()
```

```{r}
# Download the text of "The Beautiful and Damned" using its Gutenberg ID
beautiful_and_damned_text <- gutenberg_download(9830)

# View the first few lines of the text
head(beautiful_and_damned_text)

# Now you can proceed with your sentiment analysis using the downloaded text

```

### Filtering the Loughran Lexicon

Lexicon Filtering:
I filtered the Loughran sentiment lexicon to include only positive and negative words. This lexicon is used for sentiment analysis on the selected corpus.

```{r}
#filtering the loughran lexicon only for positive an negative words
loughran_posneg <- get_sentiments("loughran") %>% 
  filter(sentiment == "positive" | sentiment =="negative")
```

### Sentiment Analysis

Tokenization:
The text is tokenized into individual words, making it easier to analyze sentiment on a word-by-word basis.

```{r}
# Tokenize the text into words
beautiful_and_damned_words <- beautiful_and_damned_text %>%
  unnest_tokens(word, text)
beautiful_and_damned_words
```

Joining with Lexicon:

In this part, the tokenized words are joined with the filtered Loughran sentiment lexicon to associate sentiment scores with each word in the corpus.
```{r}
# Join with the Loughran sentiment lexicon
beautiful_and_damned_sentiments <- beautiful_and_damned_words %>%
  inner_join(loughran_posneg)

```

Sentiment Counting:

Count the occurrences of positive and negative words in the corpus to understand the overall sentiment distribution.

```{r}
# Count the occurrences of positive and negative words
sentiment_counts <- beautiful_and_damned_sentiments %>%
  count(sentiment)
```

Sentiment Visualization:

The sentiment distribution is visualized using a bar plot, where each sentiment (positive/negative) is represented along with its frequency count

```{r}
# Visualize the sentiment distribution with value counts
ggplot(sentiment_counts, aes(x = sentiment, y = n, fill = sentiment)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = n), vjust = -0.5, color = "black", size = 3) + # Add value counts
  labs(title = "Sentiment Distribution in 'The Beautiful and Damned'",
       x = "Sentiment",
       y = "Frequency") +
  theme_minimal() +
  theme(legend.position = "none")

```

### Word Clouds

Positive Sentiment Word Cloud:

I created a word cloud visualization for words associated with positive sentiment in the corpus, showing the most frequently occurring positive words.

```{r}
# Create wordclouds for positive and negative sentiments
positive_words <- beautiful_and_damned_sentiments %>%
  filter(sentiment == "positive") %>%
  count(word, sort = TRUE)
print(positive_words)

# Plot wordcloud for positive sentiment
wordcloud(positive_words$word, positive_words$n,
          max.words = 100, scale=c(3,0.5),
          colors = brewer.pal(8, "Dark2"),
          random.order = FALSE,
          rot.per = 0.35,
          main = "Wordcloud for Positive Sentiment")
```

Negative Sentiment Word Cloud:

Similarly, I created a word cloud visualization for words associated with negative sentiment in the corpus, showing the most frequently occurring negative words.

```{r}
negative_words <- beautiful_and_damned_sentiments %>%
  filter(sentiment == "negative") %>%
  count(word, sort = TRUE)
print(negative_words)

# Plot wordcloud for negative sentiment
wordcloud(negative_words$word, negative_words$n,
          max.words = 100, scale=c(3,0.5),
          colors = brewer.pal(8, "Dark2"),
          random.order = FALSE,
          rot.per = 0.35,
          main = "Wordcloud for Negative Sentiment")
```

### Conclusion

In this extended R Markdown document, I have retained the base code for sentiment analysis with Jane Austen's texts and incorporated additional analysis with "The Beautiful and Damned" by F. Scott Fitzgerald. Additionally, I have included sentiment analysis using the Loughran lexicon for "The Beautiful and Damned" text. The base code used in this R markdown document is from the "Text Mining with R" book, Chapter 2.

















